---
title: "Group Assignment (Principle of Data Science)"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(shiny)
library(rsconnect)
```

## Group members:
1. LEE YONG TAI (WQD170055)
2. VOON ZHI LING (WQD170056)
3. SOON EVONNE (WQD170057)

## Date:
10th December 2017

## Objective:
1. To estimate the monthly home loan repayment
2. To estimate how many percentage of the total payment goes into interest and principal repayment
3. To estimate the interest, principal repayment and balance after each monthly payment.

## How to use:
1. Enter your home price (in MYR)
2. Enter your downpayment amount (in MYR)
3. Slide to your interest rate (Annual Interest Rate in Percentage)
4. Slide to your housing loan period (in Years)

```{r, echo = FALSE}
shinyApp(
  
  ui = fluidPage(
  titlePanel("Home Loan Calculator"),
  
  sidebarLayout(
    
    sidebarPanel(
      #sliderInput("homeprice", "Home Price", min=10000, max=1000000, value=20000, step=100),
      textInput("homeprice", "Home Price", value=100000),
      #sliderInput("downpayment", "Down Payment", min=0, max=100000, value=0, step=100),
      textInput("downpayment", "Down Payment", value=10000),
      sliderInput("irate", "Interest Rate", min=2.00, max=10.00, value=3.25, step=0.05),
      sliderInput("loanyear", "Loan Period", min=5, max=40, value=20, step=1)
      #sliderInput("principal", "Principal", min=100, max=10000, value=500, step=100)
    ), #sidebarPanel ends
    mainPanel(
      tabsetPanel(
        tabPanel("Summary",
                 h2(p("WQD7001 Principles of Data Science : Group Assignment"), align="center"),
                 h1(strong("Home Loan Calculator"),align="center", style = "font-family: 'times'; font-si20pt"),
                 
                 h2("Group members:"),
                 p("1. LEE YONG TAI (", span(em("WQD170055"), style="color:blue"), ")", style="font-family: 'times'; font-size: 13pt"),
                 p("2. VOON ZHI LING (", span(em("WQD170056"), style="color:blue"), ")", style="font-family: 'times'; font-size: 13pt"),
                 p("3. SOON EVONNE (", span(em("WQD170057"), style="color:blue"), ")", style="font-family: 'times'; font-size: 13pt"),
                 
                 h2("Date:"),
                 p("10th December 2017", style="font-family: 'times'; font-size: 15pt"),
                 
                 h2("Objective:"),
                 p("1. To estimate the monthly home loan repayment", style="font-family: 'times'; font-size: 15pt"),
                 p("2. To estimate how many percentage of the total payment goes into interest and principal repayment", style="font-family: 'times'; font-size: 15pt"),
                 p("3. To estimate the interest, principal repayment and balance after each monthly payment", style="font-family: 'times'; font-size: 15pt"),
                 
                 h2("How to use:"),
                 p("1. Enter your ", code("Home Price"), "(", span(em("in MYR"), style="color:blue"), ")", style="font-family: 'times'; font-size: 15pt"),
                 p("2. Enter your ", code("Down Payment"), " amount (", span(em("in MYR"), style="color:blue"), ")", style="font-family: 'times'; font-size: 15pt"),
                 p("3. Slide to your ", code("Interest Rate"), "(", span(em("Annual Interest Rate in Percentage"), style="color:blue"), ")", style="font-family: 'times'; font-size: 15pt"),
                 p("4. Slide to your ", code("Loan Period"), "(", span(em("in Years"), style="color:blue"), ")", style="font-family: 'times'; font-size: 15pt")
                 ),
        tabPanel("Payment Summary", h3("Your monthly payment:",textOutput("resultPayment")), plotOutput("resultPie")),
        tabPanel("Payments Table", dataTableOutput("resultTable")))
      )
      
  ) #sidebarLayout ends
), #fluidPage ends
  
  server = function(input,output){
  #Monthly Payment Figure Output
  output$resultPayment <- renderText({
    
    initialLoan <- as.numeric(input$homeprice) - as.numeric(input$downpayment)
    balance <- initialLoan
    interestRate <- (input$irate/1200) #monthly rate
    year <- input$loanyear
    payment <- round(((balance*interestRate)/(1-(1+interestRate)^(-12*year))),2)
  }) #Monthly Payment Figure Output ends
  
  
  #Pie Chart Output
  output$resultPie <- renderPlot({
    
    initialLoan <- as.numeric(input$homeprice) - as.numeric(input$downpayment)
    interestRate <- (input$irate/1200) #monthly rate
    year <- input$loanyear
    
    firstrow <- c(0,"","","",initialLoan)
    balance <- initialLoan
    #rounded monthly payment
    payment <- round(((balance*interestRate)/(1-(1+interestRate)^(-12*year))),2)
    
    for(i in 1:((input$loanyear)*12)) {
      
      month <- as.numeric(i) #month
      interest <- round((interestRate * balance), 2) #monthly interest payment
      
      if (i != ((input$loanyear)*12)) {
        principal <- round((payment - interest), 2)
        balance <- round((balance - principal), 2)
        #interest <- (interestRate * balance) #monthly interest payment
        #principal <- (payment-interest)
        #balance <- (balance - principal)
        binded <- cbind(month, payment, interest, principal, balance)
      }      
      else {
        lastpayment <- balance + interest
        principal <- round((lastpayment - interest), 2)
        balance <- round((balance - principal), 2)
        binded <- cbind(month, lastpayment, interest, principal, balance)
      }
      
      if(exists("pie_table")!=TRUE) {
        pie_table <- cbind(month, payment, interest, principal, balance)
      }
      else {
        pie_table <- rbind(pie_table, binded) 
      }
    }
    
    total_interest <- sum(pie_table[,3]) #calculate the sum of interest paid
    total_principal <- sum(pie_table[,4]) #calculate the sum of principal paid
    pie_slice <- c(total_interest, total_principal)
    percentage_label <- c(round(pie_slice/sum(pie_slice)*100),0) #calculate the % of each slice
    pie_label <- c("Interest", "Principal") #naming the slice
    pie_label <- paste(pie_label, percentage_label) #adding the percentage value to the name of the slice
    pie_label <- paste(pie_label, "%", sep="") #adding the percentage, %, sign to the slice (without space)
    pie(pie_slice, pie_label, col=c("red", "cornsilk"), main="Payments Breakdown")
  }) #Pie Chart Output ends
  
  
  #Data Table Output
  output$resultTable <- renderDataTable({
    
    initialLoan <- as.numeric(input$homeprice) - as.numeric(input$downpayment)
    interestRate <- (input$irate/1200) #monthly rate
    year <- input$loanyear
    
    firstrow <- c(0,"","","",initialLoan)
    balance <- initialLoan
    #rounded monthly payment
    payment <- round(((balance*interestRate)/(1-(1+interestRate)^(-12*year))),2)
    
    for(i in 1:((input$loanyear)*12)) {
      
      month <- as.numeric(i) #month
      interest <- round((interestRate * balance), 2) #monthly interest payment
      
      if (i != ((input$loanyear)*12)) {
        principal <- round((payment - interest), 2)
        balance <- round((balance - principal), 2)
        #interest <- (interestRate * balance) #monthly interest payment
        #principal <- (payment-interest)
        #balance <- (balance - principal)
        binded <- cbind(month, payment, interest, principal, balance)
      }      
      else {
        lastpayment <- balance + interest
        principal <- round((lastpayment - interest), 2)
        balance <- round((balance - principal), 2)
        binded <- cbind(month, lastpayment, interest, principal, balance)
      }
      
      if(exists("complete_table")!=TRUE) {
        complete_table <- cbind(month, payment, interest, principal, balance)
      }
      else {
        complete_table <- rbind(complete_table, binded) 
      }
    }
    
    total_payments <- sum(complete_table[,2])
    total_interest <- sum(complete_table[,3])
    total_principal <- sum(complete_table[,4])
    
    tab <- cbind(total_principal,total_interest,total_payments)
    
    complete_table <- rbind(firstrow, complete_table)
    
    complete_table
    
  }) #Data Table Output ends
},
  
  options=list(width="100%", height=2000)

) #shinyApp ends
```
